### 2015-01-12 prelim/oral_prelim.tex
update to test out truncating below

\begin{frame}{Non-spatial analysis: Block maxima}
\begin{adjustwidth}{1em}{0em}
  Fisher-Tippett-Gnedenko \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n$ be i.i.d.
    \item Consider the block maximum $M_n = \max(X_1, \ldots, X_n)$
    \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
    \begin{align*}
      \frac{ M_n - b_n }{ a_n } \converged G(z)
    \end{align*}
    then $G(z)$ follows a generalized extreme value distribution (GEV) (Gnedenko, 1943)
    \item This motivates the use of the GEV for block maximum data
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Multivariate representations}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Marginally at each site, observations follow a GEV distribution
    \item For a finite collection of sites the multivariate representation for the GEV (mGEV) is
    \begin{align*}
      \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp[-V(\bz)]\\
            V(\bz)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $V(\bz)$ is called the exponent measure
      \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
      \item $H$ is a probability measure on $\Delta_d$
      \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution}
\begin{adjustwidth}{1em}{0em}
  \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
  \begin{align*}
    Y(\bs) \sim \text{skew-t}(\bX(\bs) \bbeta, \bOmega, \alpha, \text{df}=2a)
  \end{align*}
  where
  \begin{itemize} \setlength{\itemsep}{0.25em}
    \item $\bX(\bs) \bbeta$ is the location
    \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
    \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
    \item $\bSigma$ is a positive definite correlation matrix
    \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Extension of the skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Skew-$t$ distribution addresses two modeling concerns \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Extremal dependence
      \item Reasonable computing
    \end{itemize}
    \item Our contribution is to extend the skew-$t$ \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Censoring to focus on extreme observations
      \item Partitioning to address long-range dependence
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item 50 datasets for each setting \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item 144 sites in $[0, 10] \times [0, 10]$ \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
        \item 100 training
        \item 44 testing
      \end{itemize}
    \end{itemize}
    \item Model parameters \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Spatial range: $\rho = 1$
      \item Skew parameter: $\lambda = 3$
      \item Degrees of freedom: 6 for $t$ distributions
      \item Proportion of spatial variation: $\gamma = 0.9$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Simulation study results}
\centering
  \begin{figure}
    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-t1.pdf}
    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-st1.pdf} \\
    \includegraphics[width=0.3\linewidth, trim=2in 3.25in 2in 3in]{./plots/pot/bs-sim-legend.pdf}
    \caption{Relative Brier score results}
  \end{figure}
\end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Many different analysis methods incorporating \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Gaussian, symmetric-$t$, skew-$t$, and max-stable marginal distributions
      \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
      \item 4 threshold levels for $t$ marginals \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
         \item $T = 0$
         \item $T = 50$ppb, $q(0.48)$
         \item $T = 75$ppb, $q(0.92)$
         \item $T = 85$ppb, $q(0.97)$
      \end{itemize}
      \item Thresholded at $T = 75$ for max-stable
    \end{itemize}
    \item Compare Brier scores from two-fold cross validation
  \end{itemize}
\end{frame}

@@

\begin{frame}{Probability of exceedance}
  \centering
  \begin{figure}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-gaus.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-t1.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-t6.pdf}
    \caption{Probability of exceeding 75ppb at least two days for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
  \end{figure}
\end{frame}

\begin{frame}{Quantile plots}
  \centering
  \begin{figure}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-gaus.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-t1.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-t6.pdf}
    \caption{$\widehat{q}(0.95)$ for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
  \end{figure}
\end{frame}

\begin{frame}{Quantile plots}
  \centering
  \begin{figure}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-gaus.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-t1.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-t6.pdf}
    \caption{$\widehat{q}(0.99)$ for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
  \end{figure}
\end{frame}

@@

\begin{frame}{Rare spatial binary regression}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item We model $Y_i = I(Z_i > 0)$ where $Z_i\sim$ multivariate GEV (mGEV)  is a latent variable
    \item Hierarchical model for mGEV (Reich and Shaby, 2012)
    \begin{align*}
      Z(\bs) = U(\bs) \theta(\bs)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $U(\bs) \iid$ GEV(1, $\alpha$, $\alpha$) is a nugget effect
      \item $\theta(\bs) = [\sum_{l = 1}^L A_l w_l(\bs)^{1/\alpha}]^\alpha$ is the spatial process
      \item $A_l \iid$ Positive Stable($\alpha$) is a random effect representing the intensity
      \item $w_l(\bs)$ gives the weight of the intensity of the $l$th random effect on site $\bs$
      \item $\alpha \in (0, 1)$ controls strength of nugget relative to spatial dependence
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Future simulation study and data application}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Simulation study \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Data generated using logistic, Cloglog, and GEV links \vspace{0.25em}
      \begin{itemize}
        \item Exploring how rarity of event impacts prediction
      \end{itemize}
      \item Models fit using \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
        \item mGEV
        \item Random effects Gaussian distribution
      \end{itemize}
    \end{itemize}
    \item Data application: Modeling crime data \vspace{0.5em}
    \begin{itemize}
      \item Homicides, car theft, vandalism
    \end{itemize}
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{Non-spatial analysis: Block maxima}
\begin{adjustwidth}{1em}{0em}
  Fisher-Tippett-Gnedenko theorem \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n$ be i.i.d.
    \item Consider the block maximum $M_n = \max(X_1, \ldots, X_n)$
    \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
    \begin{align*}
      \frac{ M_n - b_n }{ a_n } \converged G(z)
    \end{align*}
    then $G(z)$ follows a generalized extreme value distribution (GEV) (Gnedenko, 1943)
    \item This motivates the use of the GEV for block maximum data
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Multivariate representations}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Marginally at each site, observations follow a GEV distribution
    \item For a finite collection of sites the representation for the multivariate GEV (mGEV) is
    \begin{align*}
      \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp[-V(\bz)]\\
            V(\bz)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $V(\bz)$ is called the exponent measure
      \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
      \item $H$ is a probability measure on $\Delta_d$
      \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution}
\begin{adjustwidth}{1em}{0em}
  \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
  \begin{align*}
    Y(\bs) \sim \text{skew-t}[\bX(\bs) \bbeta, \bOmega, \alpha, \text{df}=2a]
  \end{align*}
  where
  \begin{itemize} \setlength{\itemsep}{0.25em}
    \item $\bX(\bs) \bbeta$ is the location
    \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
    \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
    \item $\bSigma$ is a positive definite correlation matrix
    \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Extension of the skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Skew-$t$ distribution addresses three modeling concerns \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Flexible tails
      \item Extremal dependence
      \item Reasonable computing
    \end{itemize}
    \item Our contribution is to extend the skew-$t$ \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Censoring to focus on extreme observations
      \item Partitioning to address long-range dependence
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item 50 datasets for each setting \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item 144 sites in $[0, 10] \times [0, 10]$ \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
        \item 100 training
        \item 44 testing
      \end{itemize}
      \item 50 days
    \end{itemize}
    \item Model parameters \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Spatial range: $\rho = 1$
      \item Skew parameter: $\lambda = 3$
      \item Degrees of freedom: $2a = 6$
      \item Proportion of spatial variation: $\gamma = 0.9$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

% \begin{frame}{Simulation study results}
% \centering
%   \begin{figure}
%     \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-t1.pdf}
%     \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-st1.pdf} \\
%     \includegraphics[width=0.3\linewidth, trim=2in 3.25in 2in 3in]{./plots/pot/bs-sim-legend.pdf}
%     \caption{Relative Brier score results}
%   \end{figure}
% \end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Many different analysis methods incorporating \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Gaussian, symmetric-$t$, skew-$t$, and max-stable marginal distributions
      \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
      \item 4 threshold levels for $t$ marginals \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
         \item $T = 0$
         \item $T = 50$ ppb, $q(0.48)$
         \item $T = 75$ ppb, $q(0.92)$
         \item $T = 85$ ppb, $q(0.97)$
      \end{itemize}
      \item Thresholded at $T = 75$ for max-stable
    \end{itemize}
    \item Compare Brier scores from two-fold cross validation
  \end{itemize}
\end{frame}

@@

\begin{frame}{Probability of exceedance (2 days)}
  \centering
  \begin{figure}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-gaus.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-t1.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-t6.pdf}
    \caption{Probability of exceeding 75ppb at least two days for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
  \end{figure}
\end{frame}

\begin{frame}{Quantile plots: $\widehat{q}(0.95)$}
  \centering
  \begin{figure}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-gaus.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-t1.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-t6.pdf}
    \caption{$\widehat{q}(0.95)$ for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
  \end{figure}
\end{frame}

\begin{frame}{Quantile plots: $\widehat{q}(0.99)$}
  \centering
  \begin{figure}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-gaus.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-t1.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-t6.pdf}
    \caption{$\widehat{q}(0.99)$ for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
  \end{figure}
\end{frame}

@@

\begin{frame}{Rare spatial binary regression}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item We model $Y_i = I(Z_i > 0)$ where $Z_i\sim$ mGEV is a latent variable
    \item Hierarchical model for mGEV (Reich and Shaby, 2012)
    \begin{align*}
      Z(\bs) = U(\bs) \theta(\bs)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $U(\bs) \iid$ GEV(1, $\alpha$, $\alpha$) is a nugget effect
      \item $\theta(\bs) = [\sum_{l = 1}^L A_l w_l(\bs)^{1/\alpha}]^\alpha$ is the spatial process
      \item $A_l \iid$ Positive Stable($\alpha$) is a random effect representing the intensity
      \item $w_l(\bs)$ gives the weight of the intensity of the $l$th random effect on site $\bs$
      \item $\alpha \in (0, 1)$ controls strength of nugget relative to spatial dependence
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Future simulation study and data application}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Simulation study \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Data generated using logistic, Cloglog, and GEV links \vspace{0.25em}
      \begin{itemize}
        \item Exploring how rarity of event impacts prediction
      \end{itemize}
      \item Models fit using \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
        \item mGEV
        \item Random effects Gaussian distribution
      \end{itemize}
    \end{itemize}
    \item Data application: Modeling ozone exceedances \vspace{0.5em}
  \end{itemize}
\end{frame}
